---
title: "genes2Mmusculus"
author: "Victor Barrera"
date: "08/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(biomaRt)
library(RCurl)
library(dplyr)
library(rio)
```

# Cell Cycle

## Download latest human version

```{r}
human_cc_genes_file <- getURL("https://raw.githubusercontent.com/hbc/tinyatlas/master/cell_cycle/Homo_sapiens.csv")
human_cc_genes <- read.csv(text = human_cc_genes_file)
```

## Download Ensembl data
```{r}
ensembl <- useMart("ensembl")
ensembl.human <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")
```

## Obtain ortholog information

```{r}
ortholog_info <-getBM(attributes = c('ensembl_gene_id','mmusculus_homolog_ensembl_gene','mmusculus_homolog_orthology_type','mmusculus_homolog_orthology_confidence'),filters = 'ensembl_gene_id',values =human_cc_genes$geneID, mart = ensembl.human)
```

## Filter orthologs. 

We only keep the ones with one2one orthology type and high orthology confidence

```{r}
mouse_selected_genes <- ortholog_info %>% dplyr::filter(mmusculus_homolog_orthology_type == "ortholog_one2one",mmusculus_homolog_orthology_confidence == 1)
```

## Add CC info

```{r}
mouse_cc_genes <- mouse_selected_genes %>% dplyr::inner_join(human_cc_genes, by = c("ensembl_gene_id" = "geneID")) %>% 
    dplyr::select(phase,geneID=mmusculus_homolog_ensembl_gene) %>% dplyr::arrange(phase)
```

## Add date

```{r}
mouse_cc_genes$modified <- Sys.Date()
```

## Save file

```{r}
rio::export(mouse_cc_genes, file = file.path("..","cell_cycle","Mus_musculus.csv"))
```

# Cell type

## Download latest human version

```{r}
human_ct_genes_file <- getURL("https://raw.githubusercontent.com/hbc/tinyatlas/master/cell_type/Homo_sapiens.csv")
human_ct_genes <- read.csv(text = human_ct_genes_file)
```

## Download Ensembl data
```{r}
ensembl <- useMart("ensembl")
ensembl.human <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")
```

## Obtain ortholog information

```{r}
ortholog_info <-getBM(attributes = c('ensembl_gene_id','mmusculus_homolog_ensembl_gene','mmusculus_homolog_orthology_type','mmusculus_homolog_orthology_confidence'),filters = 'ensembl_gene_id',values =human_ct_genes$geneID, mart = ensembl.human)
```

## Filter orthologs. 

We only keep the ones with one2one orthology type and high orthology confidence

```{r}
mouse_selected_genes <- ortholog_info %>% dplyr::filter(mmusculus_homolog_orthology_type == "ortholog_one2one",mmusculus_homolog_orthology_confidence == 1)
```

## Add CC info

```{r}
mouse_ct_genes <- mouse_selected_genes %>% dplyr::inner_join(human_ct_genes, by = c("ensembl_gene_id" = "geneID")) %>% 
    dplyr::select(cellType,geneID=mmusculus_homolog_ensembl_gene,highConf,scaffold,notes) %>% dplyr::arrange(cellType)
```

## Add date

```{r}
mouse_ct_genes$modified <- Sys.Date()
```

## Save file

```{r}
rio::export(mouse_ct_genes, file = file.path("..","cell_type","Mus_musculus.csv"))
```


 